---
title: "Hemorragia SubAracnoidea - PAASH"
author: "Raul RC"
output:
  pdf_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warnings=FALSE)
```

```{r libs, echo=FALSE, message=FALSE, warning=FALSE,}
# Libraries
library(pander)
library(knitr)
library(kableExtra)
library(scales)
# set working directory
git="/home/bic/rcruces/Desktop/rcruces-Yeatman/git_here"
setwd(paste0(git,"/hsa_paash"))

# Functions
group.test <- function(X,Group){
  print(kruskal.test(X~Group))
  print(pairwise.wilcox.test(X, Group, p.adj="bonferroni", exact=F))
  aovFA <- aov(X~Group)
  print(TukeyHSD(aovFA))
}
```

# Prognosis on Admission of Aneurysmal Subarachnoid Hemorrhage (PAASH)  
PAASH scale is based on the Glasgow Comma Scale (GCS) assigned at the time of patient admission.  
```{r gcs, echo=FALSE, warning=FALSE, message=FALSE}
df <- data.frame(PAASH=c("I","II","III","IV","V"), GCS=c("15","11-14","8-10","4-7","3"))
pander(df)
```


# Clinical table by PAASH
```{r cli, echo=FALSE, warning=FALSE, message=FALSE}
hsa.cli <- read.csv("data/PAASH_clinica.csv")
colnames(hsa.cli)[1] <- " "
pander(hsa.cli)
```

# Aneurysm location:  age and gender difference in aneurysm distribution  

```{r loc, echo=FALSE, warning=FALSE, message=FALSE}
hsa.loc <- read.csv("data/PAASH_localizacion.csv")
pander(hsa.loc[,c("X","ACA","ACM","ACP","basilar","carotida","otra")])
```

# Percentage of patients with poor outcome  
The percentage was calculated based on the PAASH score.  
  
## Glassgow Outcome Scale (GOS)  
**GOS 1:** Death  
**GOS 2:** Persistent vegetative state  
**GOS 3:** Severe disaability  
**GOS 4:** Moderate disability  
**GOS 5:** Good recovery  
  
## Outcome based on GOS  
> **Poor:** GOS 1-3  
> **Fair:** GOS 4-5

There is a significant difference in the percentage of poor outcome between the original study and our cohort for PAASH-I. In our hospital patients with PAASH I have better outcome.

```{r out, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10}
hsa.out <- read.csv("data/PAASH-outcome.csv")
# Comparison between proportions X-squared
for (i in 1:5) {
 res <- prop.test(x = c(hsa.out$poor[i],hsa.out$poor[i+5]), n=c(hsa.out$N.sum[i],hsa.out$N.sum[i+5])) 
 print(paste0("PAASH-",i,", X-squared: ",round(res$statistic,2),", p-val: ",round(res$p.value,2)))
}


# Percentage of poor outcome
per <- cbind(hsa.out[,10]/hsa.out$N.sum*100, hsa.out[,10]/hsa.out$N.total*100)
par(mfrow=c(1,2))
plot(1:5, per[1:5,1], type="b", col=NA, axes = FALSE, ylab="Percentage of poor outcome", xlab="PAAHS", main="% per PAAHS")
axis(1,at=1:5,labels = hsa.out$paash[1:5])
axis(2,at=seq(0,100,20),labels = seq(0,100,20),las=2)
points(1:5, per[6:10,1],cex=2,pch=18,col="gray65"); lines(1:5, per[6:10,1],lwd=2,col="gray65")
points(1:5, per[11:15,1],cex=1.5,pch=17,col="gray65"); lines(1:5, per[11:15,1],lwd=2,col="gray65",lty=2)
points(1:5, per[1:5,1],cex=1.5,pch=19); lines(1:5, per[1:5,1],lwd=2)

# plot(1:5, per[1:5,2], type="b", col=NA, axes = FALSE, ylab="Percentage of poor outcome", xlab="PAAHS", main="% of total sample size",ylim = c(0,25))
# axis(1,at=1:5,labels = hsa.out$paash[1:5])
# axis(2,at=seq(0,100,20),labels = seq(0,100,20),las=2)
# points(1:5, per[6:10,2],cex=2,pch=18,col="gray65"); lines(1:5, per[6:10,2],lwd=2,col="gray65")
# points(1:5, per[11:15,2],cex=1.5,pch=17,col="gray65"); lines(1:5, per[11:15,2],lwd=2,col="gray65",lty=2)
# points(1:5, per[1:5,2],cex=1.5,pch=19); lines(1:5, per[1:5,2],lwd=2)

Col=c("lightgoldenrod1", "sandybrown", "indianred2", "red3", "brown4")
plot(per[1:5,1], per[6:10,1], bty='n', xlab = 'Mexican Population', ylab = 'Original study', axes = FALSE, xlim=c(0,100), ylim=c(0,100),col=NA,main="Percentage of poor outcome")
axis(1); axis(2,las=2)
abline(h=seq(0,100,20),v=seq(0,100,20),col="gray85")
abline(lty=2, col='red3', lwd=1.5, coef = c(0,1))
points(per[1:5,1], per[6:10,1],col="gray70",bg=Col,pch=21,cex=3)
```

# PAASH - GOS relation by cohort  
```{r gos, echo=FALSE, warning=FALSE, message=FALSE}
hsa <- read.csv("data/HSA_database.csv")
col.paash <- matrix(rep(alpha(Col,0.5),5),nrow = 5,ncol = 5)
col.gos <- matrix(rep(alpha(rev(Col),0.5),5),nrow = 5,ncol = 5,byrow = T)
# Distribution by total N
# tab <- prop.table(table(paste0("paash.",hsa$paash), paste0("gos.",hsa$GOS)))*100
# plot(1:5,1:5,col=NA, ylab="GOS", xlab="PAASH", axes = FALSE,xlim=c(0.5,5.5),ylim=c(0.5,5.5), main="Distribution by Total N=135")
# axis(1,at=1:5)
# axis(2,at=1:5,las=2)
# for (i in 1:5) {for (j in 1:5) {
#   if (tab[i,j] != 0) {
#     C=tab[i,j]/10
#     D="gray45"
#   } else {C=0; D="white"}
#   points(i,j, cex=C, col=col.paash[i,j],pch=19); points(i,j, cex=C, col=col.gos[i,j],pch=19)
#   text(i,j+0.45,paste0(round(tab[i,j],1),"%"),col = D)
# }}
# Distribution by PAASH
tab.1 <- table(paste0("paash.",hsa$paash), paste0("gos.",hsa$GOS))
tab.1 <- tab.1/rowSums(tab.1)*100
plot(1:5,1:5,col=NA, ylab="GOS", xlab="PAASH", axes = FALSE,xlim=c(0.5,5.5),ylim=c(0.5,5.5), main="Distribution by PAASH")
axis(1,at=1:5)
axis(2,at=1:5,las=2)
for (i in 1:5) {for (j in 1:5) {
  if (tab.1[i,j] != 0) {
    C=tab.1[i,j]/10
    D="gray45"
  } else {C=0; D="white"}
  points(i,j, cex=C, col=col.paash[i,j],pch=19); points(i,j, cex=C, col=col.gos[i,j],pch=19)
  text(i-0.5,j,paste0(round(tab.1[i,j],1),"%"),col = D)
}}
```

For each PAASH, no significant differences where found for sex or age
```{r dat, echo=FALSE, warning=FALSE, message=FALSE, size = 'scriptsize'}

# Gender & age
group.test(as.numeric(hsa$sex), as.factor(hsa$paash))
group.test(hsa$age, as.factor(hsa$paash))

# Linear discriminant Analisis for variable selection
```
